An Introduction to Rocker: Docker Containers for R

### 2017-07-19 manuscript.Rmd
WIP: a bit more text on the "extensible" section

Users can easily add additional software to any running Rocker image using
the standard R and Debian mechanisms. To make these changes persistent, Rocker 
images can easily be extended by any user writing their own Dockerfiles based
on an appropriate Rocker image.

Rocker images provide a few utility functions to streamline this process, including the 
`littler` [@littler] scripts which provide a concise syntax for installing packages in Dockerfiles, e.g.

@@

Users writing their own Dockerfiles are encouraged to follow the same practices as the
Rocker Project, such as the [Dockerfile Best Practices](), the use of automated builds,
and when appropriate, versioned tags.  Users can ensure their automated builds are rebuilt
every time the relevant upstream DOckerfile is updated by using build triggers on Docker Hub.

@@

- Dockerfiles. community input.  [`hlapp/rpopgen`](https://github.com/NESCent/popgen-docker)

@@

>>>>

Any portable computational environment faces an inevitable tension between
the "kitchen sink problem" at one extreme, and the "discovery problem" on
the other.  A kitchen sink image seeks to accomidate too many use cases
in a single image.  Such images are inevitably very large and thus slow
or difficult to deploy, maintain and optimize.  At the other extreme, providing
too many specialized images makes it more difficult for a user to discover 
the one they need.  The Rocker project seeks to avoid both of these problems
by providing images that an be easily extended by individuals and communities.

Users can easily add additional software to any running Rocker image using
the standard R and Debian mechanisms. To make these changes persistent, Rocker 
images can be extended by any user writing their own Dockerfiles based
on an appropriate Rocker image.  Many users maintain their own Dockerfiles
which add software specific to their own needs on top of an existing Rocker image.
This provides a convenient way to reap all the benefits listed above in a 
customized environment.

Sharing these Dockerfiles can also facilitate the emergence 
of extensions tuned to particular communities.  For instance, the `rocker/geospatial`
image emerged from the input of a number of Rocker users all adding common
geospatial libraries and packages on top of the existing Rocker images. 
This coalescence helped create a more fine-tuned image with broad support 
for a wide range of commonly used data formats and libraries.  Other community
images are developed and maintained independently of the Rocker Project,
such as the `popgen` image of population-genetics-oriented software developed 
by the National Evolutionary Synthesis Center (NESCent).

<!-- see docs instead -->
Rocker images provide a few utility functions to streamline this process, including the 
`littler` [@littler] scripts which provide a concise syntax for installing packages in Dockerfiles, e.g.

@@

Users writing their own Dockerfiles are encouraged to follow the same practices as the
Rocker Project, such as the Dockerfile Best Practices [@best-practices], the use of automated builds,
and when appropriate, versioned tags.  Users can ensure their automated builds are rebuilt
every time the relevant upstream DOckerfile is updated by using build triggers on Docker Hub.

@@

@@

Lessons for related efforts

Future directions, Open Container Initiative, Singularity & Shifter (support for Docker on HPC)
